---
title: "ORA, GSEA and enrichment map"
date: "`r Sys.Date()`"
package: NPATools
bibliography: references.bib
output:
  rmarkdown::html_vignette:
    toc: true
    number_sections: true
vignette: >
  %\VignetteIndexEntry{ORA, GSEA and enrichment map}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, echo = FALSE, message = FALSE, warning=FALSE}
devtools::load_all()
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  error    = FALSE,
  warning  = FALSE,
  #eval     = FALSE,
  message  = FALSE#,
  #fig.width = 7,
  #fig.height = 7
)
```

In this vignette we will show how to use perform Over-Representatin Analysis (ORA) and Gene Set Enrichment Analysis (GSEA) with NPATools.

# ORA and GSEA

ORA is performed by means of the function `ora()`, which requires a list of one or more gene vectors (`wb`), the reference universe and a list of gene sets to test (`gsl`):

```{r, include=TRUE, message=FALSE, warning=FALSE, eval=FALSE}
ora_res <- ora(wb = wb, universe = unverse, gsl = gsl)
```

GSEA is performed by means of the function `gsea()`, which requires a matrix with one or more columns of gene-related statistics (`rl`) and a list of gene sets to test (`gsl`):

```{r, include=TRUE, message=FALSE, warning=FALSE, eval=FALSE}
gsea_res <- gsea(rl=rl, gsl = gsl)
```

By default, each column of `rl` is sorted in decreasing order.

## Visualization

The functions `plot_ora_heatmap()` and `plot_gsea_heatmap()` draw heatmaps with results of ORA and GSEA:

```{r, include=TRUE, message=FALSE, warning=FALSE, eval=FALSE}
plot_ora_heatmap(ora_res, cluster_columns = F, cluster_rows = F, a = .5, na_col = "gold")
plot_gsea_heatmap(gsea_res = gsea_res, cluster_columns = F, cluster_rows = F, a = .5, na_col = "gold")
```

## Inter-operability

We provide the functions `ora2enrich()` and `gsea2enrich()` to obtain instances of "enrichResult" class of the package clusterProfiler:

```{r, include=TRUE, message=FALSE, warning=FALSE, eval=FALSE}
ora_enrich <- ora2enrich(ora_res, gsl = gsl, wb = wb, universe = unverse)
gsea_enrich <- gsea2enrich(gsea_res, rl = rl, gsl = gsl)
```

# Functional clustering and enrichment map

The following procedure calculates an enrichment map and the corresponding gene set clusters. These can be used, for example, to create visualizations where gene-associated values are clustered by gene membership in similar gene sets. This procedure is often called "functional clustering" because gene sets defines gene functions. The two required inputs are:

- gene_set_scores: a named vector of gene set scores;
- gene_set_list: a named list of gene set vectors.

The first step is the calculation of all-pairs similarities among gene sets:

```{r, include=TRUE, message=FALSE, warning=FALSE, eval=FALSE}
gene_set_sim <- calc_gs_sim(gene_set_list)
```

The function is based on `proxy::simil()` and, by default `gene_set_sim()`, uses "Simpson" similarity measure, also known as overlap coefficient; a list of all available measures can be obtained using `summary(proxy::pr_DB)` and more information about a particular measure can be shown by `proxy::pr_DB$get_entry("Simpson")` (see `?proxy::simil()` for further details). Once we have all-pairs similarities we can create the corresponding enrichment map (which requires gene set sizes):

```{r, include=TRUE, message=FALSE, warning=FALSE, eval=FALSE}
gs_size <- setNames(lengths(gene_set_list), names(gene_set_list))
em_res <- enrichment_map(gs_scores=gene_set_scores, gene_set_sim=gene_set_sim, gs_size=gs_size)
```

A relevant argument is `min_sim` (by default 0.2), which defines the minimum threshold for similarity values that will be considered (the others are set to zero). A data frame of gene sets, showing their cluster, score and size can be retrieved using:

```{r, include=TRUE, message=FALSE, warning=FALSE, eval=FALSE}
gene_set_clusters_df <- get_gene_set_clusters(em_res)
```
A named list of gene set vectors that defines gene-cluster membership can be obtained by:

```{r, include=TRUE, message=FALSE, warning=FALSE, eval=FALSE}
clust_genes <- get_cluster_genes(gene_set_clusters=gene_set_clusters_df, gene_set_list=gene_set_list)
```
